Lossy compression of DNA sequencing quality data

### 2015-04-21 manuscript/q-compression.tex
Added missing text.

We have simply provided here the initial steps in analyzing the effect
of lossy compression on quality values using a single,
high-coverage bacterial data set. More work needs to be done using
additional biological data sets, such as human and mouse, along with
different sequencing technologies. A more direct comparison against
related lossy compression tools, such as
\textsc{SlimGene}~\citep{Kozanitis:2011kl} and
\textsc{\textsc{q}ual\textsc{c}omp}~\citep{Ochoa:2013rt}, needs to be
performed. Additionally, other types of sequencing data can be
analyzed apart from the Illumina data examined here. For example, the
PacBio sequencing instruments produce very long reads (with average
read lengths on the order of 10 kbp), but with the trade-off of having
a high error-rate ($\sim$15\%). Unlike the class of quality values we
have examined here, the distribution of erroneous bases is relatively
uniform~\cite{Ferrarini:2013vf}. The assembly complexity of bacterial
genomes can be greatly simplified, producing near complete genome
assemblies, by utilizing a single run of these long
reads~\citep{Koren:2013ye}. If long read sequencing technologies such
as PacBio become more widely adopted, it would be of huge benefit to
examine the potential of lossy compression algorithms on not only the
quality values, but the biological sequencing data themselves.

>>>>

We have simply provided here the initial steps in analyzing the effect
of lossy compression on quality sequences using two
high-coverage bacterial, human, and mouse data sets.
A more direct comparison against additional lossy compression tools, such as
\textsc{SlimGene}~\citep{Kozanitis:2011kl}, needs to be
performed. Furthermore, other types of sequencing data can be
analyzed apart from the Illumina data examined here. For example, the
PacBio sequencing instruments produce very long reads (with average
read lengths on the order of 10 kbp), but with the trade-off of having
a high error-rate ($\sim$15\%). Unlike the class of quality values we
have examined here, the distribution of erroneous bases is relatively
uniform~\cite{Ferrarini:2013vf}. The assembly complexity of bacterial
genomes can be greatly simplified, producing near complete genome
assemblies, by utilizing a single run of these long
reads~\citep{Koren:2013ye}. If long read sequencing technologies such
as PacBio become more widely adopted, it would be of huge benefit to
examine the potential of lossy compression algorithms on not only the
quality values, but the biological sequencing data themselves.
